---
title: "run_admixture.Rmd"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir =
    "/Users/slacksa/OneDrive - The University of Colorado Denver")

library(tidyverse)
library(conflicted)
library(data.table)

```

# **Overview**

RMarkdown runs Admixture using 1000G as reference. This report is not intended
to be knit - just to be run chunk by chunk.ß

Requires PLINK1.9, PLINK2, bcftools, and ADMIXTURE.


# **Run Admixture**

### Steps 0-1 & 0-2

Skipping scripts 0-1 and 0-2 - already run by Shane in ancestry_estimation.qmd.

### Step 1-1

*TODO: does it make sense to use raw exome chip, not QCed one?*

```{zsh engine.opts='-i', eval = F}
mamba activate bcftools
code_dir="/Users/slacksa/repos/tools/admixture"
RKJCOLLAB="/Users/slacksa/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver"

bash "$code_dir"/1-1_get_comm_variants.sh \
  -r ${RKJCOLLAB}/Immunogenetics_T1D/genetics/teddy_r01/1000genomesPCA/qc/1000G.chr1.qc.pruned \
  -s ${RKJCOLLAB}/DAISY/genetics/daisy_ask_genetics/daisyask_exome_array/clean/daisyexome_study_ids \
  -o ${RKJCOLLAB}/DAISY/genetics/daisy_ask_genetics/admixture

```

Note: multi-allelic warning at first merge (by-chr merge of reference) -
for now, just removing these variants but could revisit in future.

17827 overlapping variants for Admixture.


### Step 1-2

```{zsh engine.opts='-i', eval = F}
mamba activate bcftools
code_dir="/Users/slacksa/repos/tools/admixture"
RKJCOLLAB="/Users/slacksa/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver"

bash "$code_dir"/1-2_merge.sh \
  -o ${RKJCOLLAB}/DAISY/genetics/daisy_ask_genetics/admixture

```


17824 variants for admixture after flipping and excluding variants that coulnd't
be flipped.

### Step 1-3

```{zsh engine.opts='-i', eval = F}
mamba activate bcftools
code_dir="/Users/slacksa/repos/tools/admixture"
RKJCOLLAB="/Users/slacksa/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver"

bash "$code_dir"/1-3_prune.sh \
  -o ${RKJCOLLAB}/DAISY/genetics/daisy_ask_genetics/admixture

```

16489 variants after pruning.

### Step 2-1

This step is slow and may be skipped if already done for reference population.
Use the K selected by CV in step 2-2.

SDS testing: K=1: 0.68715, K=2: 0.64285

```{zsh engine.opts='-i', eval = F}
mamba activate bcftools
code_dir="/Users/slacksa/repos/tools/admixture"
RKJCOLLAB="/Users/slacksa/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver"

bash "$code_dir"/2-1_admixture_cv.sh \
  -o ${RKJCOLLAB}/DAISY/genetics/daisy_ask_genetics/admixture

```


### Step 2-2


```{zsh engine.opts='-i', eval = F}
mamba activate bcftools
code_dir="/Users/slacksa/repos/tools/admixture"
RKJCOLLAB="/Users/slacksa/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver"

bash "$code_dir"/2-2_admixture.sh \
  -o ${RKJCOLLAB}/DAISY/genetics/daisy_ask_genetics/admixture \
  -k 5

```

### Plot - TEMP

*TODO: move this to report?*

```{r}

# plot the Q estimates from ADMIXTURE for the 1000G reference
Q.1000 <- as.data.frame(
  fread("DAISY/genetics/daisy_ask_genetics/admixture/ancestry_estimation/1000G_ref.5.Q"))

colnames(Q.1000) <- paste0("Ancestry", 1:ncol(Q.1000))

demo.1000g <- fread("Immunogenetics_T1D/raw/genetic_maps/1000genomes-phase3/integrated_call_samples_v3.20130502.ALL.panel") %>% 
  dplyr::select(-V5, -V6)

fam.1000g <- fread("DAISY/genetics/daisy_ask_genetics/admixture/ancestry_estimation/1000G_ref.fam") %>% 
  rename(FID = V1,
         IID = V2) %>% 
  left_join(demo.1000g, by = c("IID"="sample")) %>% 
  select("IID","super_pop")

Q.1000$IID <- fam.1000g$IID
Q.1000$pop <- fam.1000g$super_pop

Q.1000 <- Q.1000 %>% arrange(pop)
Q.1000$Individual <- factor(Q.1000$IID, levels = Q.1000$IID)

Q.1000_long <- Q.1000 %>%
  pivot_longer(
    cols = starts_with("Ancestry"),
    names_to = "AncestryComponent",
    values_to = "Proportion"
  )

# Compute midpoint positions for population labels
pop_positions <- Q.1000 %>%
  group_by(pop) %>%
  summarize(mid = mean(as.numeric(Individual)))

colors <- RColorBrewer::brewer.pal(n = 8, name = "Set2")[1:5]

# Plot stacked barplot with population labels
ggplot(Q.1000_long, aes(x = Individual, y = Proportion, fill = AncestryComponent)) +
  geom_bar(stat = "identity", width = 1) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  labs(x = "Population",
       y = "Ancestry",
       title = "1000 Genomes Reference") +
  geom_text(
    data = pop_positions,
    aes(x = mid, y = -0.05, label = pop),
    inherit.aes = FALSE
  )

Q.1000_long %>% 
  dplyr::filter(pop == "EUR") %>% 
  group_by(AncestryComponent) %>% 
  summarise(mean = mean(Proportion),
            min = min(Proportion),
            max = max(Proportion))
# lets say if Ancestry1 > 0.80


```


```{r}
# Plot the study estimates
Q.study <- as.data.frame(
  fread("DAISY/genetics/daisy_ask_genetics/admixture/ancestry_estimation/study.5.Q"))

colnames(Q.study) <- paste0("Ancestry", 1:ncol(Q.study))

fam.study <- fread("DAISY/genetics/daisy_ask_genetics/admixture/ancestry_estimation/study.fam") %>% 
  rename(FID = V1,
         IID = V2) %>% 
  select(FID, IID)

Q.study$IID <- fam.study$IID
Q.study$pop <- ifelse(Q.study$Ancestry1 > 0.80, "EUR", "other")

table(Q.study$pop)

Q.study <- Q.study %>% arrange(pop)
Q.study$Individual <- factor(Q.study$IID, levels = Q.study$IID)

Q.study_long <- Q.study %>%
  pivot_longer(
    cols = starts_with("Ancestry"),
    names_to = "AncestryComponent",
    values_to = "Proportion"
  )

# Compute midpoint positions for population labels
pop_positions <- Q.study %>%
  group_by(pop) %>%
  summarize(mid = mean(as.numeric(Individual)))

colors <- RColorBrewer::brewer.pal(n = 8, name = "Set2")[1:5]

# Plot stacked barplot with population labels
ggplot(Q.study_long, aes(x = Individual, y = Proportion, fill = AncestryComponent)) +
  geom_bar(stat = "identity", width = 1) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  labs(x = "Population",
       y = "Ancestry",
       title = "study Ancestries") +
  geom_text(
    data = pop_positions,
    aes(x = mid, y = -0.05, label = pop),
    inherit.aes = FALSE
  ) + 
  geom_hline(yintercept = 0.20)

demo.study <- Q.study %>% 
  left_join(fam.study, by = c("IID"= "IID")) %>% 
  select(FID, IID, pop)
```

```{r}
# Write out all populations results
write_tsv(demo.study,
          "DAISY/genetics/daisy_ask_genetics/admixture/ancestry_estimation/study_pops.txt")

# Write out list with > 80% EUR in DAISY
demo.study.eur <- demo.study %>%
  dplyr::filter(pop == "EUR")
write_tsv(demo.study.eur,
          "DAISY/genetics/daisy_ask_genetics/admixture/ancestry_estimation/study_eur_list.txt")

```



# **Session Info**

R Session info:

```{r session_info}
sessionInfo()
```